---
title: "Tutorial 2"
author: "Aditya Rao"
output:
  html:
    toc: true
    toc_float: true
    theme: cerulean  
  pdf: 
    toc: true
    toc_float: true
    theme: cerulean  
---

# In this section we are loading all the required libraries for the tutorial

```{r Libraries, message = FALSE, warning= FALSE, echo=TRUE}
library(rmarkdown)
library(knitr)
library(tidyverse)
library(readr)
library(kableExtra)
library(ggplot2)
library(gridExtra)
```



## In this tutorial we are looking at the Coronavirus cases detected within the Hubei area as reported in the [Lancet](https://www.thelancet.com/coronavirus) Journal website as of March 12, 2019.


# Reading the data
```{r Table data, message=FALSE, warning=FALSE, echo= FALSE}
dat <- read_csv("Data/COVID19_March12_Hubei.csv")
```

## Data in the first 10 rows

```{r reading data rows, message=FALSE, echo=TRUE, warning=FALSE}
head(dat, n=10)
```

## Glimpse of the data

```{r glimpse, message=FALSE, echo=TRUE, warning=FALSE}
glimpse(dat)
```

## Rows and columns of the data
The dataset has `r nrow(dat)` rows and `r ncol(dat)`columns in total.

### Why is it important to know the dimension of your dataset?
*1. Choosing analysis methods: Certain data analysis techniques require a specific number of dimensions to function correctly.
*2. Understanding data complexity: The number of features can influence the complexity of your data and the difficulty of extracting insights.
*3.Preventing overfitting: With high-dimensional data (many features), machine learning models can overfit the data, leading to poor performance on unseen examples. 
4. Optimizing algorithms: Knowing the dimensionality allows you to choose algorithms that are efficient for that specific size and structure of data.


## What are the variable names in the dataset? 

```{r Names of the Variables, message=FALSE, echo=TRUE, warning=FALSE}
names(dat)
```
### Explaining the 'AGE' and 'Date of Admission to Hospital' variables in the Dataset.

Age: The patient's chronological age for medical context.
Date of Admission to Hospital: The date the patient was officially checked into the hospital


## Cleaning the dataset

```{r cleaning, echo=TRUE, warning=FALSE, message=FALSE}
cleaned <-  dat |>
    select(country, age, sex, city, province, latitude, longitude)
```

* `age` is a character (`<chr>`)
* `sex` is a character (`<chr>`)
* `city` is a character  `<chr>`
* `province` is a character  `<chr>`
* `latitute` is double (numeric) `<dbl>`
* `Longitude` is double (numeric) `<dbl>`

### Converting age to numeric 
```{r converting into numeric values, message=FALSE, echo=FALSE, warning=FALSE}
cleaned <- cleaned |>
  mutate(age = as.numeric(age))
```

## Inspecting the Data
``` {r inspecting the data, message=FALSE, echo=TRUE, warning= FALSE}
head(cleaned$age, 30)

round(sum(is.na(cleaned$age)), digits = 2)
```

### Dimensions of the new dataset
```{r dimensions, warning=FALSE, echo=FALSE, message=FALSE}

dim(cleaned)
dim(cleaned)
```









